---
title: "Host-spacer networks in simulation 12499"
author: "Shai Pilosof"
date: "April 16, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_knit$set(root.dir = "/Users/Shai/GitHub/ecomplab/CRISPR_networks/")
```

```{r LOAD PACKAGES, include=FALSE}
library(tidyverse)
library(magrittr)
library(sqldf)
library(cowplot)
library(igraph)
library(bipartite)
library(infomapecology)
library(ggtree)
library(ape)
library(treeio)
```

```{r initialize, include=FALSE}
# !diagnostics off

# parameters of the model; used by the functions, so need to initialize here
beta=50;phi=10^-7;p=10^-5;q=10^-5;m=0.1;spacer_len=10

on_Midway <- function(){ifelse(Sys.getenv('USER')=='pilosofs',T,F)}

if (length(commandArgs(trailingOnly=TRUE))==0) {
  # args <- c('mu5e-7_initialDiffDp1_S50P15_R-13997','5*10^-7','15',F, F)
  args <- c('mu1e-7_initialDiffDp1_S10P15_R-12499','1*10^-7','15',F, T)
} else {
  args <- commandArgs(trailingOnly=TRUE)
}
base_name <- args[1]
mu <- eval(parse(text = args[2]))
protospacer_len <- eval(parse(text = args[3]))
make_plots <- as.logical(args[4])
complete_analysis <- as.logical(args[5])

# Functions ---------------------------------------------------------------
prep.packages <- function(package.list, verbose=T) {
  loaded = package.list %in% .packages()
  if ( all(loaded) ) return(invisible())
  
  package.list = package.list[!loaded]
  installed = package.list %in% .packages(TRUE)
  if ( !all(installed) ) install.packages(package.list[!installed], repos="http://cran.rstudio.com/")
  for ( p in package.list ){
    print(paste("Loading package:",p))
    if (verbose){
      library(p,character.only=TRUE)
    } else {
      suppressMessages(library(p,character.only=TRUE))  
    }
  }
}

returnnull <- function(x) if (is.null(x)){'none'} else {x}

notify <- function(x){
  print(paste('[',Sys.time(),'] ',x,sep=''))
}

empty <- function (web, count = FALSE) 
{
  web[is.na(web)] <- 0
  if (NCOL(web) == 1 | NROW(web) == 1) {
    if (NCOL(web) == 1 & NROW(web) != 1) {
      nr <- sum(web > 0)
      nc <- 1
    }
    if (NROW(web) == 1 & NCOL(web) != 1) {
      nc <- sum(web > 0)
      nr <- 1
    }
    if (NROW(web) == 1 & NCOL(web) == 1) {
      nr <- 1
      nc <- 1
    }
    out <- web[1:nr, 1:nc, drop = FALSE]
    if (count) 
      attr(out, "empty") <- c(`empty rows` = NROW(web) - 
                                nr, `empty columns` = NCOL(web) - nc)
    return(out)
  }
  cempty <- which(colSums(web) == 0)
  rempty <- which(rowSums(web) == 0)
  cind <- if (length(cempty) == 0) 
    1:NCOL(web)
  else (1:NCOL(web))[-cempty]
  rind <- if (length(rempty) == 0) 
    1:NROW(web)
  else (1:NROW(web))[-rempty]
  out <- web[rind, cind, drop = FALSE]
  if (count) 
    attr(out, "empty") <- c(`empty rows` = length(rempty), 
                            `empty columns` = length(cempty))
  return(out)
}

# Matrix trace
tr <- function(x){
  sum(diag(x))
}

calculate_ev_nestedness <- function(B){
  # It is faster to calculate the ev for smaller matrices. Because the leading
  # ev of BB^T and B^TB is the same, we first check how to produce A.
  if (nrow(B)<ncol(B)){
    A <- B%*%t(B)
  } else {
    A <- t(B)%*%B
  }
  ev_max <- max(eigen(A, symmetric = T, only.values = T)$values) # Not calculating eigenvectors speeds calculatoins remarkably.
  return(ev_max)
}

# The upper bound is taken from Stanić Z. Inequalities for Graph Eigenvalues.
# Cambridge University Press; 2015, page 51. It is calculated as:
# \lambda_1^2\leq \frac{tr(BB^T)}{n_1}+\sqrt{\frac{n_1-1}{n_1}\left( tr((BB^T)^2)-\frac{tr(BB^T)^2}{n_1} \right)}
# where n1 is the shorter dimension of B (also see function calculate_ev_nestedness)
calculate_upper_bound <- function(B){
  if (nrow(B)<ncol(B)){
    A <- B%*%t(B)
  } else {
    A <- t(B)%*%B
  }
  n1 <- nrow(A)
  x <- tr(A)/n1
  y <- (n1-1)/n1
  z <- tr(A%*%A)-tr(A)^2/n1
  upper_bound <- x+sqrt(y*z)
  return(upper_bound)
}

record_data <- function(x){
  notify(paste('Recording ',deparse(substitute(x)),sep=''))
  write_csv(x, paste(base_name,'_',deparse(substitute(x)),'.csv',sep=''), col_names = T)
}

make_name <- function(x,hr){
  hr_str <- str_pad(hr, width = 4, side = 'left', pad = '0')
  paste('data/',x,'_',hr_str,'.csv',sep='')
}

make_png <- function(p, method='ggsave'){
  if (method=='ggsave'){
    ggplot2::ggsave(paste('figures/',base_name,'_',deparse(substitute(p)),'.png',sep=''), p, device = 'png', width = 32, height = 18, units = 'cm', dpi = 200)
  } else {
    png(paste('figures/',base_name,'_',deparse(substitute(p)),'.png',sep=''),1920,1080,res=150)
    print(p)
    dev.off()
  }
}

make_svg <- function(p){
  svg(paste('figures/',base_name,'_',deparse(substitute(p)),'.svg',sep=''),12.8,8)
  print(p)
  dev.off()
}

make_png_svg <- function(p){
  make_png(p)
  make_svg(p)
}

standard_plot <- function(p){
  p+scale_x_continuous(breaks=label_seq)+
    geom_vline(xintercept=BDRs$start, col='#27AE60',size=1.2)+
    geom_vline(xintercept=VDRs$start, col='purple',size=1.2)+
    labs(x='Time')
}

plot_matrix <- function(M, layout='random', method = 'ggplot', binary_cols=c('gray','red'), title='', x_title='', y_title='', legend_title=''){
  M_orig <- M
  # M=M_orig
  if (layout=='random'){
    M <- M[sample(1:nrow(M), size = nrow(M), replace = F), sample(1:ncol(M), size = ncol(M), replace = F)]
  }
  
  if (layout == "diagonal") {
    ca <- cca(M)
    M <- M[order(summary(ca)$sites[, 1], decreasing = TRUE), 
           order(summary(ca)$species[, 1], decreasing = TRUE)]
  }
  
  if (layout=='nested' & method == 'ggplot'){
    tmp <- as.matrix(M[order(rowSums(M), decreasing = T), order(colSums(M),decreasing = F)])
    rnames <- rownames(M)[order(rowSums(M), decreasing = T)]
    cnames <- colnames(M)[order(colSums(M), decreasing = F)]
    rownames(tmp) <- rnames
    colnames(tmp) <- cnames
    M <- tmp
  }
  if (layout=='nested' & method == 'heatmap'){
    M <- M[order(rowSums(M), decreasing = T), order(colSums(M),decreasing = T)]
  }
  
  if (identical(unique(as.vector(M)), c(0,1))){ # If binary matrix
    colors=binary_cols
  } else {
    if (all(M==0)){ # If empty matrix
      colors='gray'
    } else {
        colors=c('gray',gg_color_hue(n=max(M))) # If neither binary nor empty
      }
  }
    
  
  if (method=='heatmap'){
    heatmap(M, Rowv = NA, Colv = NA, 
            symm = F, scale = 'none', revC = T, margins=c(7,7), 
            # labRow = F, labCol = F, 
            col=colors, main=title)
    return(invisible())
  }
  
  if (method=='ggplot'){
    M <- reshape2::melt(M)
    if (ncol(M)==1){ # If M has only one dimension after ordering M for the nestedness then reshaping does not work well.
      M$Var1 <- rownames(M)
      M$Var2 <- dimnames(M_orig)[[which(dim(M_orig)==1)]]
    }
    plt=as_tibble(M) %>% 
      ggplot()+
      geom_tile(aes(Var1,Var2,fill=value))+
      scale_fill_gradientn(colours=colors, name=legend_title)+
      theme(
        axis.text.x=element_text(angle=-90),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 18))+
      labs(title=title, x=x_title,y=y_title)+
      coord_fixed()
    return(list(M=M,plt=plt))
  }
}

gg_color_hue <- function(n, hue_min = 10, hue_max = 280, tune1 = 62, tune2 = 100) {
  hues = seq(hue_min, hue_max, length=n+1)
  hcl(h=hues, l=tune1, c=tune2)[1:n]
}

matrix_to_list_bipartite <- function(x){
  g <- graph.incidence(t(x), weighted = T)
  l_bip <- as_tibble(igraph::as_data_frame(g, "edges"))
  return(l_bip)
}

list_to_matrix <- function(l, directed=F, bipartite=T){
  # Only deal with first 3 columns.
  g <- graph.data.frame(l,directed = directed)
  if(bipartite){
    V(g)$type <- V(g)$name %in% as.data.frame(l)[,1] # As.data.frame is necessary because if l is a tibble then l[,1] does not work
    output_mat <- as_incidence_matrix(g, names = T, attr = 'w', sparse = F)
  } else {
    output_mat <- as_adjacency_matrix(g, names = T, sparse = F, attr = 'w')
  }
  # print(dim(output_mat))
  # if(any(rowSums(output_mat)==0)){stop('One or more rows sum to 0')}
  # if(any(colSums(output_mat)==0)){stop('One or more columns sum to 0')}
  return(output_mat)
}


create_networks_hr <- function(virus_data,bacteria_data,hr){
  virus_data_hr <- virus_data %>% filter(timesOfRecord==hr)
  bacteria_data_hr <- bacteria_data %>% filter(timesOfRecord==hr)
  virus_abund_hr <- virus_data_hr %>% select(label, density) %>% rename(V_ID=label) %>% mutate(V_ID=paste('V_',str_pad(V_ID, 4, 'left', '0'),sep=''))
  bacteria_abund_hr <- bacteria_data_hr %>% select(label, density) %>% rename(B_ID=label) %>% mutate(B_ID=paste('B_',str_pad(B_ID, 4, 'left', '0'),sep=''))
  
  # Create networks in a list form
  
  ## Virus-protospacer network
  virus_ps_hr_list <- virus_data_hr %>% select(-time, -timesOfRecord, -density) %>% 
    gather(key='PSidx', value='value', -label) %>% 
    rename(V_ID=label, PS=value) %>% 
    arrange(V_ID,PS) %>% 
    mutate(w=1) %>% 
    mutate(V_ID=paste('V_',str_pad(V_ID, 4, 'left', '0'),sep='')) %>% 
    mutate(PS=paste('P_',str_pad(PS, 4, 'left', '0'),sep='')) %>% 
    select(V_ID,PS,w,PSidx)
  
  ## Bacteria-spacer network
  bacteria_sp_hr_list <- bacteria_data_hr %>% 
    select(-time, -timesOfRecord, -density) %>% 
    gather(key='SPidx', value='value', -label) %>% 
    filter(value!=-1) %>%
    rename(B_ID=label, SP=value) %>% 
    arrange(B_ID,SP) %>% 
    mutate(w=1) %>%
    # mutate(w=ifelse(SP==-1,0,1)) %>% 
    mutate(B_ID=paste('B_',str_pad(B_ID, 4, 'left', '0'),sep='')) %>% 
    mutate(SP=paste('P_',str_pad(SP, 4, 'left', '0'),sep='')) %>% 
    select(B_ID,SP,w,SPidx)
 
  # Bacteria without spacers: This is needed for NEUTRAL SCENARIOS in which
  # interactions are not related to spacer-protospacer acquisition. So bacteria
  # can actually survive without acquiting protospacers.
  bacteria_no_sp <- bacteria_data_hr %>% 
    select(-time, -timesOfRecord, -density) %>% 
    gather(key='SPidx', value='value', -label) %>% 
    filter(value==-1) %>% 
    group_by(label) %>% 
    summarise(no_spacers=n()) %>% filter(no_spacers==spacer_len) %>%
    rename(B_ID=label) %>% 
    mutate(B_ID=paste('B_',str_pad(B_ID, 4, 'left', '0'),sep=''))
  
  # Add to the bacteria-spacer network those bacteria without spacers.
  suppressWarnings(
    bacteria_sp_hr_list <- bind_rows(bacteria_sp_hr_list, 
                                     as.tibble(expand.grid(B_ID=bacteria_no_sp$B_ID, SP=unique(bacteria_sp_hr_list$SP), w=0)))
  )
  ## Immunity network as a list
  x <- virus_ps_hr_list %>% select(V_ID, PS)
  y <- bacteria_sp_hr_list %>% 
    filter(w!=0) %>% # MUST include only the interactions usign filter(w!=0), otherwise the immunity network will have false edges that should not exist.
    select(B_ID, SP) 
  immunity_list <- x %>% inner_join(y, by = c('PS'='SP')) %>% 
    arrange(B_ID, V_ID) %>% 
    group_by(V_ID, B_ID) %>% count() %>%
    rename(w=n)
  
  # Define nodes in the system
  viruses_hr <- sort(unique(virus_abund_hr$V_ID))
  bacteria_hr <- sort(unique(bacteria_abund_hr$B_ID))
  spacers_hr <- sort(union(bacteria_sp_hr_list$SP, virus_ps_hr_list$PS))
  nodes <- tibble(nodeID=1:(length(viruses_hr)+length(bacteria_hr)+length(spacers_hr)),
                  nodeName=c(viruses_hr, bacteria_hr, spacers_hr),
                  type=c(rep(1,length(viruses_hr)), rep(2,length(bacteria_hr)), rep(3,length(spacers_hr)))
  )
  nodes$type <- as.integer(nodes$type)
  
  # Check networks
  if (nrow(bacteria_sp_hr_list)==0) {bacteria_sp_hr_list <- NULL}
  if (nrow(immunity_list)==0) {immunity_list <- NULL}
  
  ## Infection network
  # To get this we first need to create the immunity nertwork as a matrix
  if (!is.null(immunity_list)){
    immunity_matrix <- list_to_matrix(immunity_list) # transform to a matrix format
  } else {
    immunity_matrix <- matrix(0, nrow=length(bacteria_hr), ncol=length(viruses_hr), dimnames = list(bacteria_hr, viruses_hr))
  }

  # Add all the nodes which were not included in the list because they have degree of zero
  # missing_bacteria <- setdiff(unique(bacteria_sp_hr_list$B_ID),rownames(immunity_matrix)) 
  # immunity_matrix <- rbind(immunity_matrix, matrix(0, 
  #                                                    ncol = ncol(immunity_matrix),
  #                                                    nrow = length(missing_bacteria),
  #                                                    dimnames = list(missing_bacteria, colnames(immunity_matrix))))
  # 
  
  infection_matrix <- 1*(immunity_matrix==0) # Binary infection network is the "negative (photography-wise)" of the immunity network
  if (any(infection_matrix==1)){ # If there is at least one non-zero interaction in the infection network
    N_T <- sum(bacteria_abund_hr$density)
    # This will produce a matrix with values: (N_i*V_j)/N_T
    A <- crossprod(matrix(bacteria_abund_hr$density, nrow=1,
                          ncol=length(bacteria_abund_hr$density),
                          dimnames=list(NULL, bacteria_abund_hr$B_ID)),
                   matrix(virus_abund_hr$density, 
                          nrow=1,ncol=length(virus_abund_hr$density)                        ,
                          dimnames=list(NULL, virus_abund_hr$V_ID)))
    A <- A/N_T
    # This will remove all cells that are 0.
    A <- A[rownames(infection_matrix),colnames(infection_matrix)]
    infection_matrix <- infection_matrix*A 
    # Transform the infection network to a list
    g <- graph.incidence(t(infection_matrix), directed = F, weighted = T) # Need to transpose so the "from" would be viruses and the "to" would be bacteria
    infection_list <- as.tibble(igraph::as_data_frame(g, what = 'edges'))
    names(infection_list) <- c('V_ID', 'B_ID', 'w')
  } else { # All interactions are 0
    infection_list <- NULL
  }
  
  # Calculate the effective mutation matrix. These lines here save a double for loop
  # and selecting bacteria and virus abundances at each step.
  # First produce a matrix with N_i*V_j values:
  A <- crossprod(matrix(bacteria_abund_hr$density, nrow=1,  
                        ncol=length(bacteria_abund_hr$density),
                        dimnames=list(NULL, bacteria_abund_hr$B_ID)),
                 matrix(virus_abund_hr$density, 
                        nrow=1,ncol=length(virus_abund_hr$density)                        ,
                        dimnames=list(NULL, virus_abund_hr$V_ID)))
  A <- A[rownames(immunity_matrix),colnames(immunity_matrix)]
  M0 <- 1*(immunity_matrix==0)
  M0 <- M0*A*beta*mu*phi*(1-q) # This is from Childs et al 2012 Suppl Info, page 4. The multiplication by M0 thakes only the no matches
  M1 <- 1*(immunity_matrix!=0)
  M1 <- M1*A*beta*mu*phi*p # This is from Childs et al 2012 Suppl Info, page 4. The multiplication by M1 thakes only the  matches
  mutation_matrix <- M0+M1

  return(list(hr=hr,
              virus_ps_list=virus_ps_hr_list,
              bacteria_sp_list=bacteria_sp_hr_list,
              immunity_list=immunity_list,
              immunity_matrix=immunity_matrix,
              infection_list=infection_list,
              infection_matrix=infection_matrix,
              mutation_matrix=mutation_matrix,
              bacteria_no_sp=bacteria_no_sp,
              virus_abund_hr=virus_abund_hr,
              bacteria_abund_hr=bacteria_abund_hr,
              nodes=nodes
  ))
}

# A function to run Infomap on a unipartite/bipartite network.
run_infomap <- function(edges, remove_zero_edges = T, return_L=F){
  if (nrow(edges)==0 || is.null(edges)){
    print('No edges!! Returning NULL')
    return (NULL)
  } else {
    if (remove_zero_edges){
      edges %<>% filter(w!=0)
    }
    edges <- as.data.frame(edges[,1:3])
    # Create internal index for nodes
    Set1 <- sort(unique(edges[,1]))
    Set2 <- sort(unique(edges[,2]))
    nodes <- data.frame(nodeID=1:(length(Set1)+length(Set2)), nodeName=c(Set1,Set2))
    
    # Transform node names to IDs
    edges[,1] <- nodes$nodeID[match(edges[,1],nodes$nodeName)]
    edges[,2] <- nodes$nodeID[match(edges[,2],nodes$nodeName)]
    
    write_delim(edges, 'infomap.txt', delim = ' ', col_names = F)
    # system('./Infomap_v01926 infomap.txt . -N 20 --tree -2 --undirected --seed 123 -i link-list --silent --out-name infomap_out')
    system('./Infomap_mac_01925 infomap.txt . -N 20 --tree -2 --undirected --seed 123 -i link-list --silent --out-name infomap_out')
    modules_hr <- suppressMessages(read_delim('infomap_out.tree', delim = ' ', skip = 2, col_names = c('path', 'flow', 'name', 'node')))
    modules_hr %<>% dplyr::select(name, path, nodeID=node) %>%
      rowwise %>%
      mutate(module=str_split(path,':')[[1]][1]) %>% 
      dplyr::select(nodeID, module)
    suppressMessages(
      nodes %<>% inner_join(modules_hr) %>% mutate(module=as.integer(module), hr) %>% 
        mutate(mod_hr=paste(hr,module,sep='_')) %>%  # This is a unique identifier of the module because the number of module will repeat in other hours.
        dplyr::select(-nodeID)   # Remove that because it was an internal index for that function, so to not confuse with values outside the function
    )
    
    if (return_L){
      L <- read_lines('infomap_out.tree', n_max = 1)
      L <- parse_number(str_sub(L, str_locate(L, 'to codelength')[1], str_length(L)))
      return(list(L=L, modules=nodes))
    } else {
      return(as.tibble(nodes))
    }
  }
}



plot_modules <- function(network_list, m_df, remove_zero_edges=T){
  if (remove_zero_edges){
    network_list %<>% filter(w!=0)
  }
  M_set1 <- M_set2 <- network_list <- network_list[,1:3]
  names(M_set1) <- names(M_set2) <- names(network_list) <- c('Set1','Set2','w')
  suppressMessages(suppressWarnings(
    M_set1 %<>% left_join(m_df, by=c('Set1'='nodeName')) %>% rename(module1=module,mod_hr1=mod_hr)
  ))
  suppressMessages(suppressWarnings(
    M_set2 %<>% left_join(m_df, by=c('Set2'='nodeName')) %>% rename(module2=module,mod_hr2=mod_hr)
  ))
  suppressMessages(suppressWarnings(
    M <- full_join(M_set1, M_set2, by = c("Set1", "Set2", "w")) %>% 
      dplyr::select(hr=hr.x,Set1,Set2,w,module1,module2,mod_hr1,mod_hr2)
  ))
  Set1_modules <- unique(M_set1[,c('Set1','module1')])
  Set1_modules <- with(Set1_modules, Set1_modules[order(module1,Set1),])
  Set2_modules <- unique(M_set2[,c('Set2','module2')])
  Set2_modules <- with(Set2_modules, Set2_modules[order(module2,Set2),])
  
  # If there are no interactions outside the module then do not need the gray
  # color. Otherwise, it will plot the first module in gray.
  M %<>% mutate(edge_in_out=ifelse(module1==module2,'in','out')) %>% 
    mutate(value_mod=ifelse(edge_in_out=='in',module1,0)) %>% 
    mutate(Set1=factor(Set1, levels=Set1_modules$Set1), Set2=factor(Set2, levels=Set2_modules$Set2))
  
  module_colors <- tibble(module1=unique(M$module1), col=gg_color_hue(n=length(unique(M$module1))))
  
  suppressMessages(suppressWarnings(
  M %<>% left_join(module_colors) %>% mutate(col=ifelse(edge_in_out=='in',col,'gray'))
  ))
  
  plt <- 
    ggplot()+
    geom_tile(data=M %>% filter(w!=0), aes(Set1, Set2,fill=col), colour='black') + # Interactions within modules
     geom_tile(data=M %>% filter(w==0), aes(Set1, Set2), fill='white') + # Add nodes with no interactions, if they exist
    scale_fill_identity()+
    theme(legend.position='none',
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.text.x = element_text(angle=-90),
          axis.title = element_blank(),
          axis.ticks = element_blank())+
    coord_fixed()

  # geom_text(aes(V_ID,B_ID,label = value_mod))
  return(plt)
}

get_regimes <- function (phage_time_series, d2_threshold=0.001, do_smoothing = T, make_plots=F) {
  if (do_smoothing){print('Finding regimes with smoothing...')} else {print('Finding regimes without smoothing...')}
  # Need to use the relative abundance of phages, because the total abundaces
  # create huge numbers in the derivatives.
  x <- phage_time_series %>% 
    group_by(timesOfRecord) %>% 
    summarise(a=sum(Pdensity)) %>% 
    rename(time=timesOfRecord) %>% 
    mutate(rel_abund=a/max(a, na.rm = T))
  # Find 1st and 2nd derivatives of the time series
  d1 <- diff(x$rel_abund)/diff(x$time) 
  d1_df <- tibble(time=1:length(d1), d1=d1)
  d2 <- diff(d1_df$d1)/diff(d1_df$time) 
  d2_df <- tibble(time=1:length(d2), d2=d2)
  # Define time series points as regimes. A point is defined in a regime if its 2nd derivative is lower than a threshold, d2_threshold
  suppressMessages(
  regimes <- x %>% left_join(d1_df) %>% left_join(d2_df) %>% 
    mutate(regime=ifelse(abs(d2)<=d2_threshold,'BDR','VDR'))
  )
  # Find sequences of regimes/no regimes.
  regime_seq <- rle(regimes$regime)
  regime_seq <- tibble(len=regime_seq$lengths, regime=regime_seq$values) # len is the length of the sequence
  # This allows to see points where dynamics switch from regimes to no regimes
  regime_seq$switch=cumsum(regime_seq$len) # cumsum gives the point in the time series
  
  # Smoothing eliminates small blips where there is a few points of 'VDR'
  # surrounded by many 'BDR'. This threshold is decided by the distribution of
  # the lengths of no regime sequences. Sequences with length smaller than that
  # threshold are defined as yes.
  if (do_smoothing){
    smoothing_threshold <- quantile(subset(regime_seq, regime=='VDR')$len, 0.75)
    smoothing <- subset(regime_seq, len<=smoothing_threshold & regime=='VDR') %>% mutate(start=switch-len)
    # smoothing$start[1] <- 1
    for (i in 1:nrow(smoothing)){
      # print(smoothing$start[i]:smoothing$switch[i])
      regimes$regime[smoothing$start[i]:smoothing$switch[i]] <- 'BDR'
    }
    # After smoothing need to define regime sequences again
    regime_seq <- rle(regimes$regime)
    regime_seq <- tibble(len=regime_seq$lengths, regime=regime_seq$values)
    regime_seq$switch=cumsum(regime_seq$len)
  }
  # To be a regime, a sequences has to be larger than the largest no-regime section,
  # which is the virus outbreak with the largest duration.
  max_virus_cycle <- max(subset(regime_seq, regime=='VDR')$len)
  
  regime_end <- which(regime_seq$len>max_virus_cycle)
  regime_start <- which(regime_seq$len>max_virus_cycle)-1
  regime_end <- regime_seq[regime_end,]$switch
  regime_start <- regime_seq[regime_start,]$switch
  if (length(regime_start)<length(regime_end)){
    regime_end <- regime_end[-1]
  }
  
  regimes_df <- rbind(
    tibble(start=1, end=regime_start[1]-1, regime_type='VDR'), # Add the first no regime period
    tibble(start=regime_start, end=regime_end, regime_type='BDR'),
    tibble(start=regime_end+1, end=c(regime_start-1, stop_time)[-1], regime_type='VDR')
  )
  regimes_df$duration=regimes_df$end-regimes_df$start
  regimes_df %<>% arrange(start)
  
  # Plot some diagnostics
  if (make_plots){
    plt_diagnostics <- 
      regimes %>% 
      gather(key='key', value='value', -time, -regime) %>% 
      mutate(key=factor(key,levels = c('a','rel_abund','d1','d2'))) %>% 
      ggplot()+
      geom_line(aes(time, value))+
      geom_point(aes(time, value, color=regime), size=0.5)+
      scale_color_manual(values = c('#27AE60','purple'))+
      scale_x_continuous(breaks = seq(0,max(x$time),500))+
      # scale_x_continuous(limits=c(800,1000))+
      facet_wrap(~key, scales='free', labeller = as_labeller(c(`a` = "Virus abundance",
                                                               `d1` = "First derivative",
                                                               `d2` = "Second derivative",
                                                               `rel_abund` = "Virus relative abundance")))+
      geom_vline(xintercept=regime_start, col='#27AE60')+
      geom_vline(xintercept=regime_end, col='purple')+
      theme(legend.position = 'none')+labs(x='Time')
  } else {
    plt_diagnostics <- NULL
  }
  return(list(regimes_df=regimes_df,
              plt_diagnostics=plt_diagnostics))
}



print_pattern <- function(parent, child, parent_death, child_birth, child_death, node, i){
  sprintf("(%s:%f, %s:%f)%s",
          parent, parent_death-child_birth,
          child, child_death-child_birth,
          paste(node, i , sep = "_"))
}

nodes_dataframe_to_one_root <- function(nodes, parent, children, parent_death) {
  for(i in 1:nrow(children)) {
    newChildren <- nodes %>% filter(parent_id == children$id[i]) %>% arrange(desc(creation_time))
    if (i == 1) {
      tempParent<-parent
      parent_death <- parent_death
    }else{
      tempParent<-out
      parent_death<-children$creation_time[i-1]
    }
    if (nrow(newChildren) > 0) {
      child <- nodes_dataframe_to_one_root(nodes, children$id[i], newChildren, children$death[i])
      child_birth <- children$creation_time[i]
      child_death <- newChildren$creation_time[nrow(newChildren)]
    }else{
      child <- children$id[i]
      child_birth <- children$creation_time[i]
      child_death <- children$death[i]
    }
    #print(list(tempParent, child, parent_death, child_birth, child_death))
    out<-print_pattern(tempParent, child, parent_death, child_birth, child_death, parent, nrow(children)-i)
  } 
  return(out)
}

nodes_dataframe_to_newick <- function(nodes) {
  root <- nodes %>% filter(is.na(parent_id))
  
  stopifnot(nrow(root) == 1)
  
  children <- nodes %>% filter(parent_id == root$id) %>% arrange(desc(creation_time))
  head(children)
  out<-nodes_dataframe_to_one_root(nodes, root$id[1], children, root$death[1])
  return(paste(out,  ":", children$creation_time[nrow(children)] - root$creation_time[1], ";",sep = ""))
}



# Inititalize -------------------------------------------------------------
dir.create('figures')

virus_data <- read_delim(paste(base_name,'_data-phage.txt',sep=''), delim=' ', col_names = T)
bacteria_data <- read_delim(paste(base_name,'_data-bact.txt',sep=''), delim=' ', col_names = T)
bacteria_abundance <- read_delim(paste(base_name,'_Bacteria-abundance.txt',sep=''), delim = ' ')
phage_abundance <- read_delim(paste(base_name,'_Phage-abundance.txt',sep=''), delim = ' ')

stop_time <- min(max(virus_data$timesOfRecord), max(phage_abundance$timesOfRecord))
hr_seq <- seq(1, stop_time, 1)

print(paste('-------- Working simulation:',base_name,' | stop time: ',stop_time,' | mu: ',mu,' | protospacers: ', protospacer_len, ' | complete analysis: ',complete_analysis,' | make plots: ',make_plots,'---------'))

if (complete_analysis){
  print('Complete analysis')
} else {
  print('REDUCED analysis')
}


virus_data %<>% filter(timesOfRecord<=stop_time)
bacteria_data %<>% filter(timesOfRecord<=stop_time)
bacteria_abundance %<>% filter(timesOfRecord<=stop_time)
phage_abundance %<>% filter(timesOfRecord<=stop_time)
 
regimes_df <- get_regimes(phage_time_series = phage_abundance, do_smoothing = T)$regimes_df
if (regimes_df[nrow(regimes_df),]$duration==1){regimes_df <- regimes_df[-nrow(regimes_df),]} # remove the end of simulation spurious effect
if(on_Midway()){record_data(regimes_df)}
BDRs <- subset(regimes_df, regime_type=='BDR')
VDRs <- subset(regimes_df, regime_type=='VDR')

# Vectors with the time points of VDRs and BDRs
VDR_hrs <- unlist(apply(VDRs, MARGIN = 1, FUN = function(x) seq(x[1],x[2])))
BDR_hrs <- unlist(apply(BDRs, MARGIN = 1, FUN = function(x) seq(x[1],x[2])))

# This is for the x axis labels when plotting
label_seq <- pretty(hr_seq, n=10)
label_seq <- subset(label_seq, label_seq<stop_time)
regimes_seq <- tibble(hr=c(VDR_hrs,BDR_hrs), regime_type=c(rep('VDR',length(VDR_hrs)),rep('BDR',length(BDR_hrs)))) %>% arrange(hr)
if(on_Midway()){record_data(regimes_seq)}
```

```{r generate networks, include=FALSE}
print('Generating networks...')
all_networks <- vector(mode = 'list', length = length(hr_seq))
for (hr in hr_seq){
  nets <- create_networks_hr(virus_data, bacteria_data, hr)
  all_networks[[which(hr_seq==hr)]] <- nets
  notify(paste('Generated networks for time ',hr,'/',stop_time,sep=''))
}
```

```{r diversification, echo=FALSE}
# These data frames contain the extinction/mutation events and persistence of viruses/bacteria
virus_dynamics_list <- virus_data %>%
  filter(timesOfRecord<=max(hr_seq)) %>%
  select(timesOfRecord, label) %>%
  rename(V_ID=label) %>%
  mutate(V_ID=paste('V_',str_pad(V_ID, 4, 'left', '0'),sep=''), w=1) %>%
  arrange(timesOfRecord, V_ID)
virus_dynamics_list %<>% group_by(V_ID) %>% summarise(birth=first(timesOfRecord), death=last(timesOfRecord)) %>%
  mutate(persistence=death-birth+1)
virus_dynamics_list %<>% filter(birth>=min(hr_seq), death<=max(hr_seq))


bacteria_dynamics_list <-
  bacteria_data %>%
  filter(timesOfRecord<=max(hr_seq)) %>%
  select(timesOfRecord, label) %>%
  rename(B_ID=label) %>%
  mutate(B_ID=paste('B_',str_pad(B_ID, 4, 'left', '0'),sep=''), w=1) %>%
  arrange(timesOfRecord, B_ID)
bacteria_dynamics_list %<>% group_by(B_ID) %>% summarise(birth=first(timesOfRecord), death=last(timesOfRecord)) %>%
  mutate(persistence=death-birth+1)
bacteria_dynamics_list %<>% filter(birth>=min(hr_seq), death<=max(hr_seq))
```

```{r trees, echo=FALSE}
print(base_name)
tree_data <- read_delim(paste(base_name,'_Phage-TREE.txt',sep=''), delim = '\t',col_names = c("Recordtime","id","parent_id","creation_time"))
tree_data$id <- paste('V_',str_pad(tree_data$id, 4, 'left', '0'),sep='')
tree_data$parent_id <- paste('V_',str_pad(tree_data$parent_id, 4, 'left', '0'),sep='')
tree_data %<>% left_join(virus_dynamics_list %>% select(V_ID,death), by=c('id'='V_ID'))
tree_data[1,3] <- NA
tree_data$death[is.na(tree_data$death)]<-tree_data$creation_time[is.na(tree_data$death)]+1
tree <- nodes_dataframe_to_newick(tree_data)
writeLines(tree, 'tree.nwk')
tree_viruses <- treeio::read.tree('tree.nwk')
plt_viruses_tree <- 
    standard_plot(
      ggtree::ggtree(tree_viruses, ladderize = T) +
        ggtree::theme_tree2()
    )
plt_viruses_tree

tree_data <- read_delim(paste(base_name,'_Bacteria-TREE.txt',sep=''), delim = '\t',col_names = c("Recordtime","id","parent_id","creation_time"))
tree_data$id <- paste('B_',str_pad(tree_data$id, 4, 'left', '0'),sep='')
tree_data$parent_id <- paste('B_',str_pad(tree_data$parent_id, 4, 'left', '0'),sep='')
tree_data %<>% left_join(bacteria_dynamics_list %>% select(B_ID,death), by=c('id'='B_ID'))
tree_data[1,3] <- NA
tree_data$death[is.na(tree_data$death)]<-tree_data$creation_time[is.na(tree_data$death)]+1
tree <- nodes_dataframe_to_newick(tree_data)
writeLines(tree, 'tree_bacteria.nwk')
tree_bacteria <- treeio::read.tree('tree_bacteria.nwk')
plt_bacteria_tree <- 
      standard_plot(
        ggtree::ggtree(tree_bacteria, ladderize = F) +
          ggtree::theme_tree2()
      )
plt_bacteria_tree
```

```{r PD function}
test_PD_modules<- function(tree, module_object, node_start_letter){
  # Phylogenetic signal analysis
  D <- ape::cophenetic.phylo(tree) # Phyloegentic distance
  D <- matrix_to_list_unipartite(D, directed = T) # Use directed to make sure that the from column has all the nodes (need it for joining later)
  D <- D$edge_list
  
  # Difference between tree and matrix
  nodes_in_modules <- module_object$modules %>%
    filter(str_starts(node_name, node_start_letter)) %>%
    distinct(node_name) %>%
    mutate(node_name=str_replace_all(node_name, pattern = '\\.', ''))
  nodes_in_modules <- nodes_in_modules$node_name
  nodes_in_tree <- tree$tip.label
  # print(setdiff(nodes_in_modules, nodes_in_tree)) # In modules but not in tree
  # print(setdiff(nodes_in_tree, nodes_in_modules)) # In tree but not in modules
  # Overlapping nodes:
  overlapping <- intersect(nodes_in_tree, nodes_in_modules)
  
  # Observed modules
  M_obs <- module_object$modules %>%
    filter(str_starts(node_name, node_start_letter)) %>%
    mutate(node_name=str_replace_all(node_name, pattern = '\\.', '')) %>%
    filter(node_name %in% overlapping) %>%
    rename(m=module_level1) %>%
    select(node_name, m)
  
  #Mean PDistance between hosts within modules
  D_obs <- M_obs %>%
    inner_join(D, by=c('node_name'='from')) %>% # join PD distances
    rename(d=weight) %>%
    arrange(m, node_name) %>%
    group_by(m) %>% # Per module
    filter(to %in% node_name) %>% #Host pairs within a module
    summarise(d_mean=mean(d), mod_size=n())
  D_obs_mean <- mean(D_obs$d_mean)
  
  # print('Observed network:')
  # print(D_obs)
  
  #Shuffle to create permuted modules of the same size,
  #and recalculate the meand PD within modules. The shuffling permutes the ID of the strains.
  D_perm <- NULL
  nperm <- 500
  for (i in 1:nperm){
    # print(i)
    D_perm %<>% bind_rows(
      M_obs %>%
        mutate(node_name=sample(node_name, replace = F)) %>%
        inner_join(D, by=c('node_name'='from')) %>% # join PD distances
        rename(d=weight) %>%
        arrange(m, node_name) %>%
        group_by(m) %>% # Per module
        filter(to %in% node_name) %>% #Host pairs within a module
        summarise(d_mean=mean(d)) %>% # Calculate mean PD within modules
        mutate(run=i)
    )
  }
  
  # Null hypothesis is that the permuted distance is smaller than the observed for
  # each module (i.e., no signal). If we reject this hypothesis then there is
  # phylogenetic signal because the observed PD beteween hosts within each module
  # would be smaller than expected by chance (closely related hosts share a module).
  
  # Plot the means
  plt_across_modules <- 
    D_perm %>% group_by(run) %>%
    summarise(D_perm_mean=mean(d_mean)) %>%
    ggplot(aes(x=D_perm_mean))+geom_histogram()+geom_vline(xintercept = D_obs_mean)
  
  result_across_moduels <-
    D_perm %>% group_by(run) %>%
    summarise(D_perm=mean(d_mean)) %>%
    mutate(test=D_perm<D_obs_mean) %>%
    summarise(pvalue=sum(test)/nperm) %>%
    mutate(res=ifelse(pvalue<0.05,'Signal','No signal'))
  
  # This can also be tested per module
  plt_within_modules <-
    D_perm %>%
    full_join(D_obs, by='m') %>%
    rename(d_perm=d_mean.x, d_obs=d_mean.y) %>%
    ggplot(aes(x=d_perm))+
    geom_histogram()+
    facet_wrap(~m)+
    geom_vline(data = D_obs, aes(xintercept = d_mean))
  
  result_within_moduels <-
    D_perm %>%
    full_join(D_obs, by='m') %>%
    rename(d_perm=d_mean.x, d_obs=d_mean.y) %>%
    mutate(test=d_perm<d_obs) %>%
    group_by(m) %>%
    summarise(pvalue=sum(test)/nperm) %>%
    mutate(Signif=ifelse(pvalue<0.05,'Signal','No signal'),
           Signif_Bonferroni=ifelse(pvalue<0.05/nrow(D_obs),'Signal','No signal')) # Need to divide by number of modules for Bonferroni correction
  
  out <- list(D_obs=D_obs,
              D_obs_mean=D_obs_mean,
              plt_across_modules=plt_across_modules,
              plt_within_modules=plt_within_modules,
              result_across_moduels=result_across_moduels,
              result_within_moduels=result_within_moduels,
              nodes_in_modules=nodes_in_modules,
              nodes_in_tree=nodes_in_tree,
              overlapping=overlapping)
  return(out)
}
```

# Host-spacer networks
In this simulation there are 3 VDRs.

```{R, echo=FALSE}
VDRs
```

## Significance of host-spacer networks

We aggregate the networks within each VDR. The network for VDR 1 is not significantly modular because it si very small. The other two are significantly modular.

```{r host_spacer modularity, echo=FALSE}
hs_modularity_aggregated <- NULL
for (i in 1:nrow(VDRs)){
  metaweb <- NULL
  for (hr in VDRs$start[i]:VDRs$end[i]){
    edges <- all_networks[[which(hr_seq==hr)]]$bacteria_sp_list
    if(is.null(edges)){next}
    metaweb %<>% bind_rows(edges) %>% distinct(B_ID, SP)
  }
  metaweb %<>% mutate(w=1)
  x <- create_monolayer_object(metaweb, directed = F, bipartite = T)
  host_sp_modularity <- run_infomap_monolayer(x, infomap_executable = 'Infomap', flow_model = 'undirected', silent = T, trials = 100, two_level = T, seed = 123, signif = T, shuff_method = 'r00', nsim = 100, verbose = F)

plot_modular_matrix(host_sp_modularity)

print(paste('VDR ',i,' Pvalue for modularity: ',host_sp_modularity$pvalue,sep='')) # Is modularity significant?

tibble(L_sim=host_sp_modularity$L_sim) %>%
  ggplot(aes(L_sim))+
  geom_histogram(fill='plum')+
  geom_vline(xintercept = host_sp_modularity$L, linetype='dashed')+
  theme_bw()+
  labs(x='Map equation L', y='Count')+
  theme(legend.position='none', axis.text = element_text(size=10), axis.title = element_text(size=10))
}
```

## Phylogenetic distance in host-spacer networks

**This is done exactly as for the empirical network analysis of Russia 2010**

We shuffle the observed networks to create permuted modules of the same size and recalculate the mean PD within modules. The shuffling permutes the ID of the strains. The null hypothesis is that the permuted distance is smaller than the observed for each module (i.e., no signal). If we **reject** this hypothesis then there is phylogenetic signal because the observed PD beteween hosts within each module would be smaller than expected by chance (closely related hosts share a module). We see that if we consider the mean of mean PD across modules then we indeed reject this hypothesis, so there is a phylogenetic signal.

We see mixed results. In VDRs 1 and 2 there is no phylogenetic signal. In the third, it depends how you test (across modules or within modules) and in which modules.

```{r host_spacer phylogenetic distance, echo=FALSE}
# res <- NULL
for (i in 1:nrow(VDRs)){
  print('------------------------')
  metaweb <- NULL
  for (hr in VDRs$start[i]:VDRs$end[i]){
    edges <- all_networks[[which(hr_seq==hr)]]$bacteria_sp_list
    if(is.null(edges)){next}
    metaweb %<>% bind_rows(edges) %>% distinct(B_ID, SP)
  }
  metaweb %<>% mutate(w=1)
  x <- create_monolayer_object(metaweb, directed = F, bipartite = T)
  host_sp_modularity <- run_infomap_monolayer(x, infomap_executable = 'Infomap', flow_model = 'undirected', silent = T, trials = 1, two_level = T, seed = 123, signif = F, verbose = F)

  # Phylogenetic signal analysis
  tree <- tree_bacteria
  D <- ape::cophenetic.phylo(tree) # Phyloegentic distance
  D <- matrix_to_list_unipartite(D, directed = T) # Use directed to make sure that the from column has all the nodes (need it for joining later)
  D <- D$edge_list

  # Difference between tree and matrix
  nodes_in_modules <- host_sp_modularity$modules %>%
    filter(str_starts(node_name, 'B')) %>%
    distinct(node_name) %>%
    mutate(node_name=str_replace_all(node_name, pattern = '\\.', ''))
  nodes_in_modules <- nodes_in_modules$node_name
  nodes_in_tree <- tree$tip.label
  # print(setdiff(nodes_in_modules, nodes_in_tree)) # In modules but not in tree
  # print(setdiff(nodes_in_tree, nodes_in_modules)) # In tree but not in modules
  # Overlapping nodes:
  overlapping <- intersect(nodes_in_tree, nodes_in_modules)

  # Observed modules
  M_obs <- host_sp_modularity$modules %>%
    filter(str_starts(node_name, 'B')) %>%
    mutate(node_name=str_replace_all(node_name, pattern = '\\.', '')) %>%
    filter(node_name %in% overlapping) %>%
    rename(m=module_level1) %>%
    select(node_name, m)

  #Mean PDistance between hosts within modules
  D_obs <- M_obs %>%
    inner_join(D, by=c('node_name'='from')) %>% # join PD distances
    rename(d=weight) %>%
    arrange(m, node_name) %>%
    group_by(m) %>% # Per module
    filter(to %in% node_name) %>% #Host pairs within a module
    summarise(d_mean=mean(d), mod_size=n())
  D_obs_mean <- mean(D_obs$d_mean)
  
  print('Observed network:')
  print(D_obs)
  
  # res[[i]]$D_obs <- D_obs
  # res[[i]]$D_obs_mean <- D_obs_mean
  
  #Shuffle to create permuted modules of the same size,
  #and recalculate the meand PD within modules. The shuffling permutes the ID of the strains.
  D_perm <- NULL
  nperm <- 500
  for (i in 1:nperm){
    # print(i)
    D_perm %<>% bind_rows(
      M_obs %>%
        mutate(node_name=sample(node_name, replace = F)) %>%
        inner_join(D, by=c('node_name'='from')) %>% # join PD distances
        rename(d=weight) %>%
        arrange(m, node_name) %>%
        group_by(m) %>% # Per module
        filter(to %in% node_name) %>% #Host pairs within a module
        summarise(d_mean=mean(d)) %>% # Calculate mean PD within modules
        mutate(run=i)
    )
  }
  # res[[i]]$D_perm <- D_perm
  # print(D_perm)

  # Null hypothesis is that the permuted distance is smaller than the observed for
  # each module (i.e., no signal). If we reject this hypothesis then there is
  # phylogenetic signal because the observed PD beteween hosts within each module
  # would be smaller than expected by chance (closely related hosts share a module).

  # Plot the means
  print(
  D_perm %>% group_by(run) %>%
    summarise(D_perm_mean=mean(d_mean)) %>%
    ggplot(aes(x=D_perm_mean))+geom_histogram()+geom_vline(xintercept = D_obs_mean)
  )
  print(
  D_perm %>% group_by(run) %>%
    summarise(D_perm=mean(d_mean)) %>%
    mutate(test=D_perm<D_obs_mean) %>%
    summarise(pvalue=sum(test)/nperm) %>%
    mutate(res=ifelse(pvalue<0.05,'Signal','No signal'))
  )
  # This can also be tested per module
  
  print(
  D_perm %>%
    full_join(D_obs, by='m') %>%
    rename(d_perm=d_mean.x, d_obs=d_mean.y) %>%
    ggplot(aes(x=d_perm))+
    geom_histogram()+
    facet_wrap(~m)+
    geom_vline(data = D_obs, aes(xintercept = d_mean))
  )
  print(
  D_perm %>%
    full_join(D_obs, by='m') %>%
    rename(d_perm=d_mean.x, d_obs=d_mean.y) %>%
    mutate(test=d_perm<d_obs) %>%
    group_by(m) %>%
    summarise(pvalue=sum(test)/nperm) %>%
    mutate(Signif=ifelse(pvalue<0.05,'Signal','No signal'),
           Signif_Bonferroni=ifelse(pvalue<0.05/nrow(D_obs),'Signal','No signal')) # Need to divide by number of modules for Bonferroni correction
  )
}
```

# Infection networks

## Significance of modularity

We test for significance at the end of each VDR, after the infection network has been built. The three infection networks are significantly modular.

```{r infection networks, echo=FALSE}
# Test at the end of each VDR
infection_modularity <- NULL
PD_results=list()
for (i in 1:nrow(VDRs)){
  print('----------------------------------------')
  hr <- VDRs$end[i]
  edges <- all_networks[[which(hr_seq==hr)]]$infection_list
  x <- create_monolayer_object(edges, directed = F, bipartite = T)
  test <- run_infomap_monolayer(x, infomap_executable = 'Infomap', flow_model = 'undirected', silent = T, trials = 100, two_level = T, seed = 123, signif = T, shuff_method = 'r00', nsim = 100, verbose=F)
  infection_modularity %<>%
    bind_rows(tibble(hr=hr,
                     pvalue=test$pvalue,
                     n_hosts=ncol(x$mat),
                     n_spacer=nrow(x$mat),
                     n_interactions=sum(x$mat>0),
                     n_modules=test$m))

  print(plot_modular_matrix(test))

  tibble(L_sim=test$L_sim) %>%
    ggplot(aes(L_sim))+
    geom_histogram(fill='plum')+
    geom_vline(xintercept = test$L, linetype='dashed')+
    theme_bw()+
    labs(x='Map equation L', y='Count')+
    theme(legend.position='none', axis.text = element_text(size=10), axis.title = element_text(size=10))
  
  PD_results[[i]] <- test_PD_modules(tree_bacteria, test, 'B')
  print(PD_results[[i]]$result_across_moduels)
  print(PD_results[[i]]$result_within_moduels)
}

print(infection_modularity)
```

